A Deeper Analysis into the Trends and Factors of the US's Declining Birth Rates

### 2024-02-13 paper/paper.qmd
update figures in document

```{r graph2, fig.cap="Trends in Birth Rates by Population Subgroup", echo=FALSE}
#| echo: false
#| message: false
#| warning: false
#| label: fig-2

# replicate figure 2a from the original paper
fig_2a_data_long %>% ggplot(aes(x = year, y = Birth_Rate, color = Age_Group)) +
  geom_line(size = 1) +
  labs(x = "Year", y = "Births per 1,000 women in relevant population subgroup", color = "Age Group", caption = "Five-year age group") +
  scale_y_continuous(limits = c(0, 140), breaks = seq(0, 150, by = 20)) +
  scale_x_continuous(limits = c(1980, 2020), breaks = seq(1980, 2020, by = 5)) +
  geom_text(aes(x = 1985, y = 120), label = "Age 25-29", color = "black") +
  geom_text(aes(x = 1985, y = 100), label = "Age 20-24", color = "black") +
  geom_text(aes(x = 1985, y = 80), label = "Age 30-34", color = "black") +
  geom_text(aes(x = 1985, y = 58), label = "Age 15-19", color = "black") +
  geom_text(aes(x = 1985, y = 33), label = "Age 35-39", color = "black") +
  geom_text(aes(x = 1985, y = 10), label = "Age 40-44", color = "black") +
  theme_classic() +
  theme(legend.position = "none")

# replicate figure 2d from the original paper
fig_2d_data_long %>% ggplot(aes(x = year, y = Birth_Rate, color = Education_Level)) +
  geom_line(size = 1) +
  labs(x = "Year", y = "Births per 1,000 women in relevant population subgroup", color = "Education Level", caption = "Mother's level of education (ages 20-44)") +
  scale_y_continuous(limits = c(0, 140), breaks = seq(0, 150, by = 20)) +
  scale_x_continuous(limits = c(1990, 2020), breaks = seq(1990, 2020, by = 5)) +
  geom_text(aes(x = 1993, y = 110), label = "No high school degree", color = "black", hjust = 0) +
  geom_text(aes(x = 1993, y = 82), label = "College graduate", color = "#f8766d", hjust = 0) +
  theme_classic() +
  geom_text(aes(x = 1993, y = 62), label = "High school graduate", color = "#00bfc4", hjust = 0) +
  geom_text(aes(x = 1993, y = 45), label = "Some college", color = "black", hjust = 0) +
  theme_classic() +
  theme(legend.position = "none")

# replicate figure 2f from the original paper
merged_data_long %>% ggplot(aes(x = year, y = round(Birth_Rate, digits = 1), color = Parity)) +
  geom_line(size = 1, stat = "summary", fun = mean) +
  labs(x = "Year", y = "Births per 1,000 women in relevant population subgroup", color = "Parity", caption = "Parity (ages 15-44)") +
  scale_y_continuous(limits = c(0, 40), breaks = seq(0, 40, by = 10)) +
  scale_x_continuous(limits = c(1990, 2020), breaks = seq(1990, 2020, by = 5)) +
  geom_text(aes(x = 1993, y = 28), label = "First birth", color = "black", hjust = 0) +
  geom_text(aes(x = 1993, y = 23), label = "Second birth", color = "black", hjust = 0) +
  geom_text(aes(x = 1993, y = 12.5), label = "Third Birth", color = "black", hjust = 0) +
  geom_text(aes(x = 1993, y = 8.5), label = "Fourth+ birth", color = "black", hjust = 0) +
  theme_classic() +
  theme(legend.position = "none")
```

>>>>

```{r graph2a, fig.cap="Trends in Birth Rates by Five-year Age Group", echo=FALSE}
#| echo: false
#| message: false
#| warning: false
#| label: fig-2

# replicate figure 2a from the original paper
fig_2a_data_long %>% ggplot(aes(x = year, y = Birth_Rate, color = Age_Group)) +
  geom_line(size = 1) +
  labs(x = "Year", y = "Births per 1,000 women in \n relevant population subgroup", color = "Age Group") +
  scale_y_continuous(limits = c(0, 140), breaks = seq(0, 150, by = 20)) +
  scale_x_continuous(limits = c(1980, 2020), breaks = seq(1980, 2020, by = 5)) +
  geom_text(aes(x = 1985, y = 120), label = "Age 25-29", color = "black") +
  geom_text(aes(x = 1985, y = 100), label = "Age 20-24", color = "black") +
  geom_text(aes(x = 1985, y = 80), label = "Age 30-34", color = "black") +
  geom_text(aes(x = 1985, y = 58), label = "Age 15-19", color = "black") +
  geom_text(aes(x = 1985, y = 33), label = "Age 35-39", color = "black") +
  geom_text(aes(x = 1985, y = 10), label = "Age 40-44", color = "black") +
  theme_classic() +
  theme(legend.position = "none")
```

```{r graph2b, fig.cap="Trends in Birth Rates by Mother's Level of Education (ages 20-44)", echo=FALSE}
# replicate figure 2d from the original paper
fig_2d_data_long %>% ggplot(aes(x = year, y = Birth_Rate, color = Education_Level)) +
  geom_line(size = 1) +
  labs(x = "Year", y = "Births per 1,000 women in \n relevant population subgroup", color = "Education Level") +
  scale_y_continuous(limits = c(0, 140), breaks = seq(0, 150, by = 20)) +
  scale_x_continuous(limits = c(1990, 2020), breaks = seq(1990, 2020, by = 5)) +
  geom_text(aes(x = 1993, y = 112), label = "No high school degree", color = "black", hjust = 0) +
  geom_text(aes(x = 1993, y = 82), label = "College graduate", color = "#f8766d", hjust = 0) +
  theme_classic() +
  geom_text(aes(x = 1993, y = 62), label = "High school graduate", color = "#00bfc4", hjust = 0) +
  geom_text(aes(x = 1993, y = 45), label = "Some college", color = "black", hjust = 0) +
  theme_classic() +
  theme(legend.position = "none")
```
```{r graph2c, fig.cap="Trends in Birth Rates by Parity (ages 15-44)", echo=FALSE}

# replicate figure 2f from the original paper
merged_data_long %>% ggplot(aes(x = year, y = round(Birth_Rate, digits = 1), color = Parity)) +
  geom_line(size = 1, stat = "summary", fun = mean) +
  labs(x = "Year", y = "Births per 1,000 women in \n relevant population subgroup", color = "Parity") +
  scale_y_continuous(limits = c(0, 40), breaks = seq(0, 40, by = 10)) +
  scale_x_continuous(limits = c(1990, 2020), breaks = seq(1990, 2020, by = 5)) +
  geom_text(aes(x = 1993, y = 28), label = "First birth", color = "black", hjust = 0) +
  geom_text(aes(x = 1993, y = 23), label = "Second birth", color = "black", hjust = 0) +
  geom_text(aes(x = 1993, y = 12.5), label = "Third Birth", color = "black", hjust = 0) +
  geom_text(aes(x = 1993, y = 8.5), label = "Fourth+ birth", color = "black", hjust = 0) +
  theme_classic() +
  theme(legend.position = "none")
```
